\TITLE

### 2021-01-27 vonLaszewski-covid.tex
Updates from Overleaf

% spacing
% outcomment this for production
\usepackage{setspace}
\usepackage{lineno}
\linenumbers
%\doublespace
% above needs to be commented out for production

@@

It has an input gate, output
gate, and a forget gate. The cell is maintaining how a subsequent
value is calculated. This includes (a) how input values are
influencing the cell via the input gate, (b) how the forget gate
influences a memory value within the cell via the forget gate, and (c)
how the output gate influences the output activation of the LSTM
unit. A cell has a number of inputs and outputs that are
weighted. During a training step, these weights are learned to be
reused in a prediction process between input and output sequences.
The LSTM network will have a number of input values represented as  vector x and a number of output values in the last layer of the network predicting a number of future values \citep{lstm-explained}. For our experiments this includes the numbers of days we use as input to our prediction and the number of days we use as output of our prediction for a single cell. In our implementation the number for inputs and outputs can be adjusted. We repeat this prediction over all input sequences to create the output forecast sequences. Next we list the equations (\ref{eq:first}) - (\ref{eq:last}) used in LSTM.

\newcommand{\RH}{\mathbb{R}^{H}}
\newcommand{\RD}{\mathbb{R}^{D}}
\begin{align}
    f_t &= \sigma_g(W_f x_t + U_f h_{t-1} + b_f) \label{eq:first} \\
    i_t &= \sigma_g(W_i x_t + U_i h_{t-1} + b_i) \\
    o_t &= \sigma_g(W_o x_t + U_o h_{t-1} + b_o) \\
    \tilde{c}_t &= \sigma_c(W_c x_t + U_c h_{t-1} + b_c) \\
    c_t &= f_t \circ c_{t-1} +i_t \circ \tilde{c}_t  \\
    h_t &= o_t \circ \sigma_h(c_t) \label{eq:last}
\end{align}

We use in the euqations $\circ$ as the Hadamard product,
$\sigma_g$ is the sigmoid function for the gate,
$D$ the number of days as input features, and
$H$ the number of hidden layers.
Inputs and outputs are defined by 
$x_{t} \in \RD$ the input vector to the LSTM unit, and 
$h_{t}\in \RH$ the hidden state vector e.g., the last of which is the output vector of the LSTM. In our case the output contains the days we predict based on the input features. Other variables include
$f_t \in \RH$ is the forget gate's activation vector, 
$i_t \in \RH$ is the input/update gates activation vector, 
$o_{t}\in \RH$ output gate's activation vector, 
$\tilde{c}_{t}\in \RH$ the cell input activation vector
$c_{t}\in \RH$ the cell state vector,  
$\sigma_c$ is the cell activation function, 
$W\in \mathbb {R} ^{H\times D}$, $U\in \mathbb {R} ^{H\times H}$ and $b\in \RH$ being the weight matrices and bias vector parameters which are learned during training.
In particular we denote $W_{q}$ and $U_{q}$ contain the weights of the
input and recurrent connections, where the subscript denotes either
the input gate $i$, the output gate $o$, the forget gate $f$ or the
memory cell $c$. For a more detailed discussion about LSTM's we refer to \cite{lstm-explained}.

@@

    \begin{footnotesize}
    \centering
    \resizebox{1.0\textwidth}{!}{ 
    \begin{tabular}{|ll|lll|ll|ll|}
    \toprule 
     \rothead{Layer} & \rothead{Type}   &
     \rothead{Output Shape} &  \rothead{Parameters for 0 risk factor}  
                  &  \rothead{Parameters for 1 risk factor} &
     \rothead{Output Shape} &  \rothead{Parameters for 33 risk factors} &
     \rothead{Output Shape} &  \rothead{Parameters for 37 risk factors}\\
    \midrule
    dense & Dense    & (None,1,32) & 96   & 128  &  (None,5,32) & 1152 & (None,5,32) & 1280 \\
    LSTM & LSTM      & (None,1,16) & 3136 & 3136 &  (None,5,16) & 3136 & (None,5,16) & 3136 \\
    LSTM 1 & LSTM   & (None,16)  & 2112  & 2112 &  (None,16)   & 2112 & (None,16)   & 2112 \\
    Dense 1 & Dense & (None,32)  &  544  & 544  &  (None,32)   &  544 & (None,32)   &  544 \\
    Dense 2 & Dense & (None,30)  &  990  & 990  &  (None,30)   &  990 & (None,30)   &  990 \\
    \bottomrule
    \end{tabular}
    } ~\\
    The LSTM contains a total of 6,910 trainable parameters in
    case we include one risk factor. In the case, we do not use any
    risk factor it is 6878, and when we include 33 risk factors, it is
    7934.
    \end{footnotesize}
\end{table}

@@

\begin{figure}[!p]
    \centering
    \includegraphics[width=1.0\textwidth]{images/boxwhisker/boxplot_casesNEW2.pdf}
    \vspace{-1cm}
    \caption{Influence of risk factors on the accuracy of the
      prediction of COVID-19 cases. The y-axis represents the
      cumulative error over all input data for the cities.}
    \label{fig:box-cases}
    \bigskip

    \includegraphics[width=1.0\textwidth]{images/boxwhisker/boxplot_deathNEW2.pdf}
    \vspace{-1cm}
    \caption{Influence of risk factors on the accuracy of the
      prediction of COVID-19 deaths. The y-axis represents the
      cumulative error over all input data for the cities.}
    \label{fig:box-death}
\end{figure}

Further, the obtained sorted order of the best prediction models by
minimum error based on the risk factors for cases and deaths are as
follows:

@@

\end{description}

@@

1	&	POP\_DENSITY\_2010	&	0.055443	&	7.94	&	5	 \\
2	&	PHLTH	&	0.054795	&	8.01	&	3	 \\
3	&	BLACK	&	0.055028	&	8.12	&	4	 \\
4	&	HISPANIC	&	0.056061	&	8.22	&	5	 \\
5	&	NHOSP	&	0.054621	&	8.26	&	3	 \\
6	&	INSURANCE	&	0.054557	&	8.34	&	3	 \\
7	&	CHOLSCREEN	&	0.054712	&	8.37	&	3	 \\
8	&	DIABETES	&	0.054586	&	8.4	&	5	 \\
9	&	STROKE	&	0.05416	&	8.45	&	4	 \\
10	&	CHD	&	0.054027	&	8.46	&	5	 \\
11	&	CHECKUP	&	0.054927	&	8.51	&	5	 \\
12	&	NBEDS	&	0.054146	&	8.54	&	5	 \\
13	&	SVI\_OVERALL	&	0.054795	&	8.58	&	5	 \\
14	&	KIDNEY	&	0.0546	&	8.61	&	5	 \\
15	&	CASTHMA	&	0.055089	&	8.63	&	5	 \\
16	&	BPMED	&	0.054685	&	8.74	&	3	 \\
17	&	LPA	&	0.054323	&	8.75	&	5	 \\
18	&	CSMOKING	&	0.053716	&	8.75	&	5	 \\
19	&	ARTHRITIS	&	0.054053	&	8.83	&	3	 \\
20	&	POVERTY	&	0.055029	&	8.84	&	3	 \\
21	&	ESTBEDS	&	0.054545	&	8.86	&	4	 \\
22	&	MHLTH	&	0.054864	&	8.9	&	3	 \\
23	&	COPD	&	0.055589	&	8.93	&	4	 \\
24	&	SENIOR	&	0.055079	&	8.93	&	3	 \\
25	&	CANCER	&	0.0545	&	8.99	&	3	 \\
26	&	NBEDS/1000	&	0.054397	&	9.05	&	3	 \\
27	&	BPHIGH	&	0.054609	&	9.13	&	4	 \\
\rowcolor{black!5} 28	&	NONE	&	0.054868	&	9.26	&	3	 \\
29	&	HIGHCHOL	&	0.056429	&	9.28	&	3	 \\
30	&	BINGE	&	0.054317	&	9.32	&	5	 \\
31	&	OBESITY	&	0.054469	&	9.5	&	3	 \\
32	&	SVI\_MINORITY	&	0.055094	&	9.54	&	4	 \\
\bottomrule
\end{tabular}
%}
\end{table}

\begin{table}[!p]
\caption{Best models for deaths for each risk factor and their respective
  errors. The model that uses no risk factor is highlighted in grey.
  We present the Root Mean Square Error (RMSE), the cumulative error and how many days as input were used for this corresponding model.
  }
\label{tab:top-1-deaths}
\bigskip
\centering
%\resizebox{1.0\textwidth}{!}{ 
\begin{tabular}{rrrrrlrl}
\toprule
 \makecell{Place/\\Rank} & 
\makecell{Risk Factor} &  
\makecell{RMSE\\Deaths} &  
\makecell[r]{Minimum\\Cumu-\\lative \\Error for \\Deaths by\\Risk Factor} &  
 \makecell[r]{Number of \\Days as \\Input} \\
\midrule
1	&	SVI\_OVERALL	&	0.067457	&	18.75	&	5	 \\
2	&	CSMOKING	&	0.066898	&	19.01	&	5	 \\
3	&	HISPANIC	&	0.067489	&	19.04	&	5	 \\
\rowcolor{black!5} 4	&	NONE	&	0.06696	&	19.09	&	4	 \\
5	&	POP\_DENSITY\_2010	&	0.066968	&	19.13	&	4	 \\
6	&	SENIOR	&	0.068498	&	19.27	&	3	 \\
7	&	CHOLSCREEN	&	0.067859	&	19.43	&	5	 \\
8	&	CASTHMA	&	0.066952	&	19.46	&	5	 \\
9	&	INSURANCE	&	0.067127	&	19.48	&	5	 \\
10	&	MHLTH	&	0.068742	&	19.53	&	3	 \\
11	&	LPA	&	0.067149	&	19.54	&	5	 \\
12	&	POVERTY	&	0.067267	&	19.57	&	5	 \\
13	&	OBESITY	&	0.067123	&	19.68	&	5	 \\
14	&	CANCER	&	0.068363	&	19.70	&	3	 \\
15	&	NBEDS/1000	&	0.06752	&	19.73	&	5	 \\
16	&	BLACK	&	0.068352	&	19.83	&	3	 \\
17	&	DIABETES	&	0.067636	&	19.84	&	5	 \\
18	&	ARTHRITIS	&	0.066927	&	19.94	&	4	 \\
19	&	NBEDS	&	0.066879	&	19.95	&	4	 \\
20	&	NHOSP	&	0.067391	&	19.95	&	5	 \\
21	&	BPHIGH	&	0.067799	&	19.99	&	5	 \\
22	&	PHLTH	&	0.068737	&	20.02	&	3	 \\
23	&	CHECKUP	&	0.067543	&	20.02	&	5	 \\
24	&	CHD	&	0.06874	&	20.15	&	3	 \\
25	&	STROKE	&	0.066906	&	20.16	&	4	 \\
26	&	HIGHCHOL	&	0.068557	&	20.35	&	3	 \\
27	&	ESTBEDS	&	0.067052	&	20.42	&	4	 \\
28	&	KIDNEY	&	0.068948	&	20.49	&	3	 \\
29	&	BINGE	&	0.067447	&	20.56	&	5	 \\
30	&	BPMED	&	0.067399	&	20.69	&	5	 \\
31	&	COPD	&	0.06745	&	20.98	&	5	 \\
32	&	SVI\_MINORITY	&	0.067443	&	21.53	&	4	 \\
\bottomrule
\end{tabular}
%}
\end{table}

@@

\singlespace

\input{response-v3-content}

>>>>

% spacing
% outcomment this for production
\usepackage{setspace}
\usepackage{lineno}
\linenumbers
\doublespace
% above needs to be commented out for production

@@

\newcommand{\RH}{\mathbb{R}^{H}}
\newcommand{\RD}{\mathbb{R}^{D}}

It has an input gate, output
gate, and a forget gate. The cell is maintaining how a subsequent
value is calculated. This includes (a) how input values are
influencing the cell via the input gate, (b) how the forget gate
influences a memory value within the cell via the forget gate, and (c)
how the output gate influences the output activation of the LSTM
unit. A cell has a number of inputs and outputs that are
weighted. During a training step, these weights are learned to be
reused in a prediction process between input and output sequences.
The LSTM network will have a number of input values represented as vector x and a number of output values in the last layer of the network predicting a number of future values \citep{lstm-explained}. For our experiments this includes the numbers of days we use as input to our prediction and the number of days we use as output of our prediction for a single cell. In our implementation the number for inputs and outputs can be adjusted. We repeat this prediction over all input sequences to create the output forecast sequences. Next we list the equations (\ref{eq:first}) - (\ref{eq:last}) used in LSTM.  We use in the equations $\circ$ as the Hadamard product,
$\sigma_g$ is the sigmoid function for the gate.
With 
$D$ being the number of days as input features, and
$H$ being the number of hidden layers. 
The inputs and outputs are defined by  
$x_{t} \in \RD$ the input vector to the LSTM unit, and 
$h_{t} \in \RH$ the hidden state vector e.g., the last of which is the output vector of the LSTM. In our case the output contains the days we predict based on the input features. The equations are

\begin{align}
    f_t &= \sigma_g(W_f x_t + U_f h_{t-1} + b_f), \label{eq:first} \\
    i_t &= \sigma_g(W_i x_t + U_i h_{t-1} + b_i), \\
    o_t &= \sigma_g(W_o x_t + U_o h_{t-1} + b_o), \\
    \tilde{c}_t &= \sigma_c(W_c x_t + U_c h_{t-1} + b_c), \\
    c_t &= f_t \circ c_{t-1} +i_t \circ \tilde{c}_t,  \\
    h_t &= o_t \circ \sigma_h(c_t) \label{eq:last},
\end{align}

where 
$f_t \in \RH$ is the forget gate's activation vector, 
$i_t \in \RH$ is the input/update gates activation vector, 
$o_{t}\in \RH$ output gate's activation vector, 
$\tilde{c}_{t}\in \RH$ the cell input activation vector
$c_{t}\in \RH$ the cell state vector,  
$\sigma_c$ is the cell activation function, 
$W\in \mathbb {R} ^{H\times D}$, $U\in \mathbb {R} ^{H\times H}$ and $b\in \RH$ being the weight matrices and bias vector parameters which are learned during training.
In particular we denote $W_{q}$ and $U_{q}$ contain the weights of the
input and recurrent connections, where the subscript denotes either
the input gate $i$, the output gate $o$, the forget gate $f$ or the
memory cell $c$. For a more detailed discussion about LSTM's we refer to \cite{lstm-explained}.

@@

    \begin{footnotesize}
    \centering
    \resizebox{1.0\textwidth}{!}{ 
    %\begin{tabular}{|ll|lll|ll|ll|}
    \begin{tabular}{lllllllll}
    \toprule 
     \rothead{Layer} & \rothead{Type}   &
     \rothead{Output Shape} &  \rothead{Parameters for 0 risk factor}  
                  &  \rothead{Parameters for 1 risk factor} &
     \rothead{Output Shape} &  \rothead{Parameters for 33 risk factors} &
     \rothead{Output Shape} &  \rothead{Parameters for 37 risk factors}\\
    \midrule
    dense & Dense    & (None,1,32) & 96   & 128  &  (None,5,32) & 1152 & (None,5,32) & 1280 \\
    LSTM & LSTM      & (None,1,16) & 3136 & 3136 &  (None,5,16) & 3136 & (None,5,16) & 3136 \\
    LSTM 1 & LSTM   & (None,16)  & 2112  & 2112 &  (None,16)   & 2112 & (None,16)   & 2112 \\
    Dense 1 & Dense & (None,32)  &  544  & 544  &  (None,32)   &  544 & (None,32)   &  544 \\
    Dense 2 & Dense & (None,30)  &  990  & 990  &  (None,30)   &  990 & (None,30)   &  990 \\
    \bottomrule
    \end{tabular}
    } ~\\
    The LSTM contains a total of 6,910 trainable parameters in
    case we include one risk factor. In the case, we do not use any
    risk factor it is 6878, and when we include 33 risk factors, it is
    7934.
    \end{footnotesize}
\end{table}

@@

\begin{figure}[!p]
    \centering
    \includegraphics[width=1.0\textwidth]{images/boxwhisker/boxplot_casesNEW2.pdf}
    \vspace{-1cm}
    \caption{Influence of risk factors on the accuracy of the
      prediction of COVID-19 cases. The y-axis represents the
      cumulative error over all input data for the cities. The x-axis labels correspond to the order of models for minimum model error.}
    \label{fig:box-cases}
    \bigskip

    \includegraphics[width=1.0\textwidth]{images/boxwhisker/boxplot_deathNEW2.pdf}
    \vspace{-1cm}
    \caption{Influence of risk factors on the accuracy of the
      prediction of COVID-19 deaths. The y-axis represents the
      cumulative error over all input data for the cities. The x-axis labels correspond to the order of models for minimum model error.}
    \label{fig:box-death}
\end{figure}

\begin{comment}
Further, the obtained sorted order of the best prediction models by
minimum error based on the risk factors for cases and deaths are as
follows:

@@

\end{description}
\end{comment}

@@

1	&	POP\_DENSITY\_2010	&	0.0554	&	7.94	&	5	 \\
2	&	PHLTH	            &	0.0548	&	8.01	&	3	 \\
3	&	BLACK	            &	0.0550	&	8.12	&	4	 \\
4	&	HISPANIC	        &	0.0561	&	8.22	&	5	 \\
5	&	NHOSP	            &	0.0546	&	8.26	&	3	 \\
6	&	INSURANCE	        &	0.0546	&	8.34	&	3	 \\
7	&	CHOLSCREEN	        &	0.0547	&	8.37	&	3	 \\
8	&	DIABETES	        &	0.0546	&	8.4	&	5	 \\
9	&	STROKE	            &	0.0542	&	8.45	&	4	 \\
10	&	CHD	                &	0.0540	&	8.46	&	5	 \\
11	&	CHECKUP	            &	0.0549	&	8.51	&	5	 \\
12	&	NBEDS	            &	0.0541	&	8.54	&	5	 \\
13	&	SVI\_OVERALL	    &	0.0548	&	8.58	&	5	 \\
14	&	KIDNEY	            &	0.0546	&	8.61	&	5	 \\
15	&	CASTHMA	            &	0.0551	&	8.63	&	5	 \\
16	&	BPMED	            &	0.0547	&	8.74	&	3	 \\
17	&	LPA	                &	0.0543	&	8.75	&	5	 \\
18	&	CSMOKING	        &	0.0537	&	8.75	&	5	 \\
19	&	ARTHRITIS	        &	0.0541	&	8.83	&	3	 \\
20	&	POVERTY	            &	0.0550	&	8.84	&	3	 \\
21	&	ESTBEDS	            &	0.0545	&	8.86	&	4	 \\
22	&	MHLTH	            &	0.0549	&	8.9	&	3	 \\
23	&	COPD	            &	0.0556	&	8.93	&	4	 \\
24	&	SENIOR	            &	0.0551	&	8.93	&	3	 \\
25	&	CANCER	            &	0.0545	&	8.99	&	3	 \\
26	&	NBEDS/1000	        &	0.0544	&	9.05	&	3	 \\
27	&	BPHIGH	            &	0.0546	&	9.13	&	4	 \\
\rowcolor{black!5} 28 &	NONE	&	0.0549	&	9.26	&	3	 \\
29	&	HIGHCHOL	        &	0.0564	&	9.28	&	3	 \\
30	&	BINGE	            &	0.0543	&	9.32	&	5	 \\
31	&	OBESITY	            &	0.0545	&	9.5	&	3	 \\
32	&	SVI\_MINORITY	    &	0.0551	&	9.54	&	4	 \\
\bottomrule
\end{tabular}
%}
\end{table}

\begin{table}[!p]
\caption{Best models for deaths for each risk factor and their respective
  errors. The model that uses no risk factor is highlighted in grey.
  We present the Root Mean Square Error (RMSE), the cumulative error and how many days as input were used for this corresponding model.
  }
\label{tab:top-1-deaths}
\bigskip
\centering
%\resizebox{1.0\textwidth}{!}{ 
\begin{tabular}{rrrrrlrl}
\toprule
 \makecell{Place/\\Rank} & 
\makecell{Risk Factor} &  
\makecell{RMSE\\Deaths} &  
\makecell[r]{Minimum\\Cumu-\\lative \\Error for \\Deaths by\\Risk Factor} &  
 \makecell[r]{Number of \\Days as \\Input} \\
\midrule
1	&	SVI\_OVERALL	&	0.0675	&	18.75	&	5	 \\
2	&	CSMOKING	    &	0.0669	&	19.01	&	5	 \\
3	&	HISPANIC	    &	0.0676	&	19.04	&	5	 \\
\rowcolor{black!5} 4	&	NONE	&	0.06696	&	19.09	&	4	 \\
5	&	POP\_DENSITY\_2010	&	0.0670	&	19.13	&	4	 \\
6	&	SENIOR	        &	0.0685	&	19.27	&	3	 \\
7	&	CHOLSCREEN	    &	0.0679	&	19.43	&	5	 \\
8	&	CASTHMA	        &	0.0670	&	19.46	&	5	 \\
9	&	INSURANCE	    &	0.0671	&	19.48	&	5	 \\
10	&	MHLTH	        &	0.0687	&	19.53	&	3	 \\
11	&	LPA	            &	0.0671	&	19.54	&	5	 \\
12	&	POVERTY	        &	0.0673	&	19.57	&	5	 \\
13	&	OBESITY	        &	0.0671	&	19.68	&	5	 \\
14	&	CANCER	        &	0.0687	&	19.70	&	3	 \\
15	&	NBEDS/1000	    &	0.0675	&	19.73	&	5	 \\
16	&	BLACK	        &	0.0684	&	19.83	&	3	 \\
17	&	DIABETES	    &	0.0676	&	19.84	&	5	 \\
18	&	ARTHRITIS	    &	0.0669	&	19.94	&	4	 \\
19	&	NBEDS	        &	0.0669	&	19.95	&	4	 \\
20	&	NHOSP	        &	0.0674	&	19.95	&	5	 \\
21	&	BPHIGH	        &	0.0678	&	19.99	&	5	 \\
22	&	PHLTH	        &	0.0687	&	20.02	&	3	 \\
23	&	CHECKUP	        &	0.0675	&	20.02	&	5	 \\
24	&	CHD	            &	0.0687	&	20.15	&	3	 \\
25	&	STROKE	        &	0.0669	&	20.16	&	4	 \\
26	&	HIGHCHOL	    &	0.0686	&	20.35	&	3	 \\
27	&	ESTBEDS	        &	0.0671	&	20.42	&	4	 \\
28	&	KIDNEY	        &	0.0689	&	20.49	&	3	 \\
29	&	BINGE	        &	0.0674	&	20.56	&	5	 \\
30	&	BPMED	        &	0.0674	&	20.69	&	5	 \\
31	&	COPD	        &	0.0675	&	20.98	&	5	 \\
32	&	SVI\_MINORITY	&	0.0674	&	21.53	&	4	 \\
\bottomrule
\end{tabular}
%}
\end{table}

@@

%\clearpage
%\singlespace
%\maketitle
%\input{response-v3-content}

\clearpage
\input{response-v4-content}
